---
title: "Metabolic_Thermal_Perf_Curves"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Brief INTRO


Simulate some data for running and testing code. The data will have a quadratic structure in the simulation.
```{r}
#simulate random data
set.seed(40)
x1 <- runif(40, min = 20, max = 34)
epsilon <- rnorm(40, mean = 0, sd = 1)
y <- x1^2 + epsilon

#put it into a dataframe
raw_data <- data.frame(x1, y)

#Plot raw data
library(ggplot2)

raw_plot <- ggplot(raw_data, aes(x1, y)) + geom_point() + theme_classic() + xlab('Temperature') +      ylab('Metabolic Rate')

raw_plot
```


#Maximum Likelihood Approaches
Using maximum likelihood approaches, we can get fit models and get an idea of how well they fit the data. Using the optim() function, we can optimize the parameters of each function for the data.

```{r}
#Maximum likelihood approach using a linear model

Linear_max <- function(p, X, Y) {
  b0 <- p[1]; b1 <- p[2]; sd <- p[3]
  
  rsd <- (b0 + b1 * X ) - Y
  
  -sum(dnorm(rsd, mean = 0, sd = sd, log = T)) 
}

linear_optim <- optim(c(1, 0, 5), Linear_max, X = x1, Y = y)
linear_optim
```

```{r}
#Maximum likelihood approach for the quadratic function

Quadratic_max <- function(p, X, Y) {
  a <- p[1]; b <- p[2]; c <- p[3]
  
  rsd2 <- (a * (X^2) + b * X + c) - Y
  
  -sum(dexp(rsd2, rate = X/Y, log = T))
}

quadratic_optim <- optim(c(1, 1, 1), Quadratic_max, X = x1, Y = y)
quadratic_optim
```

```{r}
#Maximum likelihood approach for the Gaussian function

Gaussian_max <- function(p, X, Y) {
  a <- p[1]; b <- p[2]; c <- p[3]; sd <- p[4]
  
  rsd3 <- a * (exp( ((-0.5) * ((abs(X - b)) / c )^2))) - Y
  
  -sum(dnorm(rsd3, mean = 0, sd = sd, log = T)) 
}

gaussian_optim <- optim(c(0, 0, 0, 1), Gaussian_max, X = x1, Y = y)
gaussian_optim
```
```{r}
#Maximum likelihood approach for the modified Gaussian function

modGaussian_max <- function(p, X, Y) {
  a <- p[1]; b <- p[2]; c <- p[3]; d <- p[4]; sd <- p[5]
  
  rsd4 <- a * (exp( ((-0.5) * ((abs(X - b)) / c )^d))) - Y
  
  -sum(dnorm(rsd4, mean = 0, sd = sd, log = T)) 
}

modgaussian_optim <- optim(c(0, 0, 0, 2, 1), modGaussian_max, X = x1, Y = y)
modgaussian_optim
```
```{r}
#Maximum likelihood approach for the Weibull function

Weibull_max <- function(p, X, Y) {
  a <- p[1]; b <- p[2]; c <- p[3]; d <- p[4]; sd <- p[5]
  
  rsd5 <- ((a * (((d - 1) / d)^((1 - d) / d)) * (((X - b) / c) + ((d - 1) / d)^(1 / d))^(d - 1) *     (exp(-(((X - b) / c) + ((d - 1) / d)^(1 / d))^d)) ) + ((d-1) / d) ) - Y
  
  -sum(dnorm(rsd5, mean = 0, sd = sd, log = T)) 
}

weibull_optim <- optim(c(1, 1, 1, 2, 1), Weibull_max, X = x1, Y = y)
weibull_optim
```

Collecting the log-likelihood values and numbers of parameters into a vector:

```{r}
#Likelihood values (mod_LLs)
mod_LLs <- c(linear_optim$value, quadratic_optim$value, gaussian_optim$value, modgaussian_optim$value, weibull_optim$value)

#Number of model parameters for each model above (mod_pars)
mod_pars <- c(length(linear_optim$par), length(quadratic_optim$par), length(gaussian_optim$par), length(modgaussian_optim$par), length(weibull_optim$par))
```


Calculate AIC values to take into account not only likelihood ratio, but also the complexity
(number of parameters) in the model:

```{r}
mod_AICs <- (2 * log(mod_LLs)) + 2 * (mod_pars)

mod_AICs
```

#Graphing the data with the models fit
We can look at the two best models (linear and quadratic) with the data. First, set up the functions with the parameters estimated by the MLE.

```{r}
#Linear Model
linear_eq <- function(x) (linear_optim$par[2] * x) + linear_optim$par[1]

#Quadratic Model
quadratic_eq <- function(x) (quadratic_optim$par[1] * (x^2)) + (quadratic_optim$par[2] * x) +   quadratic_optim$par[3] 
```


Now, layer them onto the plot with the raw data.

```{r}
model_plot <- raw_plot + stat_function(fun = linear_eq, color = 'blue') + stat_function(fun = quadratic_eq, color = 'red') 

model_plot 
```


